Geometrization of characters of tori over non-Archimedean local fields

### 2013-08-13 shanghai.tex
Finished the proof of functoriality.
With this done, I'm tracing the origin of functoriality back to the
source. Now I wonder if G |--> GC(G) is a stack. I may be spinning out
of control here . . .

@@

@@

\section{Products}

\begin{proposition}\label{prop:product}
The rule $\boxtimes : (\cs{L}_1,\cs{L}_2)\to \cs{L}_1\boxtimes\cs{L}_2$ defines an equivalence of categories 
\[
\GC(G_1)\times(G_2) \to \GC(G_1\times G_2)
\]
and 
\[
\begin{tikzcd}[column sep=60]
\arrow{d}{t_{G_1} \times t_{G_2}} \GCiso{G_1}\times \GCiso{G_2} \arrow{r}{\boxtimes} & \arrow{d}{t_{G_1\times G_2}} \GCiso{G_1\times G_2}\\
(G_1)(\Fq)^*\times (G_2)(\Fq)^* \arrow{r}{(\chi_1,\chi_2)\mapsto \chi_1\otimes \chi_2}  & (G_1\times G_2)(\Fq)^*
\end{tikzcd}
\]
commutes.
\end{proposition}

\begin{proof}
\end{proof}

@@

%\input{snake}
\begin{theorem}\label{thm:snake}
  If $G$ is a smooth abelian group scheme, locally of finite type over
  $\Fq$, and the abelian group $\pi_0(G)(\bFq)$ (the geometric
  component group) is finitely generated, then the trace of Frobenius
  homomorphism
  \[
  \trFrob{G} : \GCiso{G} \to G(\Fq)^*
  \]
  from the Picard group of the category of geometrized characters of $G$ to
  the group of characters of $G(\Fq)$, is an isomorphism.
\end{theorem}

@@

We have already seen that the trace of Frobenius defines a canonical homomorphism $\GCiso{\GN{T}} \to \Hom_\text{ad}(T(K),\EEx)$. 
In this section we show that this homomorphism is functorial, in the following sense.

@@

\begin{proof}
When combined with the definition of $\GC(\GN{T})$ and Property~\ref{T(K)}, the first part of Proposition~\ref{prop:pullback} shows that $F_1$ is a functor, while the second part shows that Trace of Frobenius is a natural transformation $t: F_1 \to F_2$. When further combined with Proposition~\ref{prop:product}, we see that $F_1$ is an additive functor and $t: F_1 \to F_2$ is a natural transformation between additive functors.
\end{proof}

>>>>

\Clifton{Now that I'm working on the functoriality of trace of Frobenius, I  think the source of the functoriality is mainly this: $\GC$ is a stack. I'm thinking about this now.}

@@

\Clifton{Address functoriality here?}

@@

%\section{Products} 
We end this section with a simple result on products of geochars, which we will be needed to show that the isomorphism above is functorial.

\begin{lemma}\label{lemma:product}
The rule $\boxtimes : (\cs{L}_1,\cs{L}_2)\to \cs{L}_1\boxtimes\cs{L}_2$ defines an equivalence of categories 
\[
\GC(G_1)\times(G_2) \to \GC(G_1\times G_2)
\]
and 
\[
\begin{tikzcd}[column sep=60]
\arrow{d}{t_{G_1} \times t_{G_2}} \GCiso{G_1}\times \GCiso{G_2} \arrow{r}{\boxtimes} & \arrow{d}{t_{G_1\times G_2}} \GCiso{G_1\times G_2}\\
(G_1)(\Fq)^*\times (G_2)(\Fq)^* \arrow{r}{(\chi_1,\chi_2)\mapsto \chi_1\otimes \chi_2}  & (G_1\times G_2)(\Fq)^*
\end{tikzcd}
\]
commutes.
\end{lemma}

\begin{proof}
The only non-trivial part is essential surjectivity. Set $G \ceq G_1\times G_1$. Let $\cs{L}$ be a geochar on $G$. Define $f : G\to G\times G$ by $f(g_1,g_2) = (g_1,e_2,e_1,g_2)$. Observe that $m\circ f = \id_G$. 
Then the isomorphism $\mu : m^* \gcs{L} \to \gcs{L}\boxtimes \gcs{L}$ gives an isomorphism
\[
f^*\mu : f^* m^* \gcs{L} \to f^*(\gcs{L}\boxtimes \gcs{L}) = f^*p_1^*\gcs{L}\otimes f^* p_2^*\gcs{L}.
\]
Observe that $p_1\circ f = r_1 \circ q_1$ and $f_2 = r_2\circ f$ where $p_1 , p_2 : G\times G \to G$ are the usual things and $r_1 : G_1\times G_2 \to G_1$ and $r_2 : G_1\times G_2 \to G_2$ are the obvious things and $q_1 : G_1\to G_1\times G_2$ and  $q_2 : G_2\to G_1\times G_2$ are the usual things. Then $f^*p_1^*\gcs{L} = r_1^* q_1^* \gcs{L}$ and $f^*p_2^*\gcs{L} = r_2^* q_2^* \gcs{L}$. Now $\cs{L}_1 \ceq r_1^* \cs{L}$ is a geochar on $G_1$ and $\cs{L}_2 \ceq r_2^* \cs{L}$ is a geochar on $G_2$, by Proposition~\ref{prop:pullback}. Then
$ f^*p_1^*\gcs{L}\otimes f^* p_2^*\gcs{L} = r_1^*\gcs{L}_1\otimes r_2^* \gcs{L}_2 = \gcs{L}_1\boxtimes \gcs{L}_2$. In this way, $\mu$ determines an isomorphism
\[
\gamma \ceq f^*\mu : \gcs{L} \to  \gcs{L}_1\boxtimes \gcs{L}_2.
\]
So $\boxtimes$ is essentially surjective.
\end{proof}

@@

%\input{snake}
\begin{theorem}\label{thm:snake}
  If $G$ is a smooth abelian group scheme, locally of finite type over
  $\Fq$, and the abelian group $\pi_0(G)(\bFq)$ (the geometric
  component group) is finitely generated, then the trace of Frobenius
  homomorphism
  \[
  \trFrob{G} : \GCiso{G} \to G(\Fq)^*
  \]
  from the Picard group of the category of geometrized characters of $G$ to
  the group of characters of $G(\Fq)$, is an isomorphism.
  \Clifton{Address functoriality, here?}
\end{theorem}

@@

We have already seen that the trace of Frobenius defines a canonical homomorphism $\GCiso{\GN{T}} \to \Hom_\text{ad}(T(K),\EEx)$. 
In this section we show that this homomorphism is functorial, in the following sense.
\Clifton{Fold this section into the main result on characters of local tori?}

@@

\begin{proof}
When combined with the definition of $\GC(\GN{T})$ and Property~\ref{T(K)}, the first part of Proposition~\ref{prop:pullback} shows that $F_1$ is a functor, while the second part shows that Trace of Frobenius is a natural transformation $t: F_1 \to F_2$. When further combined with Lemma~\ref{lemma:product}, we see that $F_1$ is an additive functor and $t: F_1 \to F_2$ is a natural transformation between additive functors.
\end{proof}
